Selective enhancement of an exploited species undermines community stability

### 2021-11-30 rmd/manuscript.Rmd
edit: manuscript

## Maintext

@@

Our model predicted a destabilizing effect of selective enhancement on ecological communities. Increasing numbers of released individuals $R$ led to greater temporal fluctuations of communities as illustrated by increased CV (Figure XX). This pattern stemmed mainly from the reduced means of enhanced and unenhanced species groups (Figure XX) that are driven by different mechanisms. For enhanced species, released individuals induced strong overcompensation, where the negative effect of reduced per-capita reproduction due to competition exceeds the numerical benefit of released individuals (increased reproducing individuals). Therefore, the enhanced species reached the lower long-term average of population density with higher numbers of released individuals. Meanwhile, the reduced mean of unenhanced species is caused by interspecific competition between enhanced and unenhanced species. This negative effect of released individuals through competition was noticeable even with weak interspecific competition that would otherwise allow stable coexistence. Combined, the long-term average of the total community density decreased more sharply than individual species groups with increased numbers of released individuals (XX). The SDs showed a similar trend, but the relationship was flatter than the mean at the community level. Since a CV is a ratio of an SD to a mean, the steeper relationship of the mean community density has led to the increased CV.

The phenomena of the decreased long-term means should arise as long as humans control the number of released individuals (i.e., donor-control). The released individuals are free from any density dependent feedback in natural communities and therefore undermine the internal regulation of self-sustaining communities.

These results were qualitatively consistent across ecological scenarios (XX).

The desta

@@

### Methods

### Theory

@@

### Empirical analysis

#### Data

@@

#### Statistical analysis

@@

## Data availability

## Acknowledgments

## References

>>>>

# Maintext

@@

Our model predicted a destabilizing effect of selective enhancement on ecological communities. Increasing numbers of released individuals $R$ led to greater temporal fluctuations of communities as illustrated by increased CV (Figure XX). This pattern stemmed mainly from the reduced means of enhanced and unenhanced species groups (Figure XX) that are driven by different mechanisms. For enhanced species, released individuals induced strong overcompensation, where the negative effect of reduced per-capita reproduction due to competition exceeds the numerical benefit of released individuals (increased reproducing individuals). Therefore, the enhanced species reached the lower long-term average of population density with higher numbers of released individuals. Meanwhile, the reduced mean of unenhanced species is caused by interspecific competition between enhanced and unenhanced species. Combined, the long-term average of the total community density decreased more sharply than individual species groups with increased numbers of released individuals (XX). The SDs showed a similar trend, but the relationship was flatter than the mean at the community level. Since a CV is a ratio of an SD to a mean, the steeper relationship of the mean community density has led to the increased CV. These patterns were qualitatively similar across ecological scenarios.

Importantly, the reduced mean of community density may arise as long as humans control the number of releases. The stability of ecological communities rests on internal density regulation that allows for stable coexistence of competing species. For example, significant niche difference in resource use ($\alpha_{ij, i \ne j} < 1$) enables competing species to grow from low population densities because the numerically dominant species undergoes stronger intraspecific competition. However, the number of release is not subject to density-dependent feedback. The excessive individuals released by humans impose competition that would otherwise not exist and undermine the self-regulation process of stable communities. This argument is supported by the fact that the destabilizing effect of selective enhancement manifested even with weak interspecific competition that allows for stable coexistence of competing species. Perhaps, the only exception would be a scenario in which the species of interest is on the verge of local extinction with little or no density regulation (i.e., very large carrying capacity).

@@

# Methods

## Theory

@@

## Empirical analysis

### Data

@@

### Statistical analysis

@@

# Data availability

# Acknowledgments

# References
